---
title: "In-class Exercise 5"
author: "Zhang Cunlei"
date: "16 Dec 2023"
date-modified: "last-modified"
format: html
execute: 
  echo: true
  eval: true
  warning: false
editor: visual
---

## Getting Started

```{r}
#| eval: false
devtools::install_github("LuKeCe/spflow")
```

Next, we will load spflow and other R packages into R.

```{r}
pacman::p_load(tmap, sf, spdep, sp, Matrix, spflow, reshape2, knitr, tidyverse)
```

## Preparing spflow objects

First, we will load the data.

```{r}
mpsz_nb <- read_rds("data/rds/mpsz_nb.rds")
mpsz_flow <- read_rds("data/rds/mpsz_flow.rds")
mpsz_var <- read_rds("data/rds/mpsz_var.rds")
```

### Creating `spflow_network-class` objects

For our model, we choose the contiguity based neighborhood structure.

```{r}
mpsz_net <- spflow_network(
  id_net = "sg",
  node_neighborhood = 
nb2mat(mpsz_nb$by_contiguity),
  node_data = mpsz_var,
  node_key_column = "SZ_CODE")

mpsz_net
```

```{r}
mpsz_net_pairs <- spflow_network_pair(
  id_orig_net = "sg",
  id_dest_net = "sg",
  pair_data = mpsz_flow,
  orig_key_column = "ORIGIN_SZ",
  dest_key_column = "DESTIN_SZ")

mpsz_net_pairs
```

Next, we will combine `mpsz_net` and `mpsz_net_pairs` .

```{r}
mpsz_multi_net <- spflow_network_multi(mpsz_net, mpsz_net_pairs)

mpsz_multi_net
```

### Correlation Analysis

This step aims to avoid including explanatory variables that are highly correlated. The code chunk below will be used to show the correlation matrix.

```{r}
cor_formula <- log(1 + TRIPS) ~
  BUSSTOP_COUNT +
  AGE7_12 +
  AGE13_24 +
  AGE25_64 +
  SCHOOL_COUNT +
  BUSINESS_COUNT +
  RETAILS_COUNT +
  FINSERV_COUNT +
  P_(log(DISTANCE + 1))

cor_mat <- pair_cor(
  mpsz_multi_net,
  spflow_formula = cor_formula,
  add_lags_x = FALSE)

colnames(cor_mat) <- paste0(
  substr(
    colnames(cor_mat),1,3),"...")

cor_image(cor_mat)
```

## Model Calibration

#### The base model

The code chunk will be like

```{r}
base_model <- spflow(
  spflow_formula = log(1 + TRIPS) ~
    O_(BUSSTOP_COUNT +
         AGE25_64) +
    D_(SCHOOL_COUNT+
    BUSINESS_COUNT +
    RETAILS_COUNT +
    FINSERV_COUNT) +
    P_(log(DISTANCE + 1)),
  spflow_networks = mpsz_multi_net)

base_model
```

#### Residual diagnostics

Next, `pair_cor()` will be used to inspect the relationship of the residual and the explanatory variables by using the code chunk below.

```{r}
corr_residual <- pair_cor(base_model)
colnames(corr_residual) <- substr(colnames(corr_residual),1,3)
cor_image(corr_residual)
```

```{r}
spflow_formula <- log(1 + TRIPS) ~
  O_(BUSSTOP_COUNT +
       AGE25_64) +
  D_(SCHOOL_COUNT+
  BUSINESS_COUNT +
  RETAILS_COUNT +
  FINSERV_COUNT) +
  P_(log(DISTANCE + 1))

model_control <- spflow_control(
  estimation_method = "mle",
  model = "model_1")

mle_model <- spflow(
  spflow_formula,
  spflow_networks = mpsz_multi_net,
  estimation_control = model_control)

mle_model
```
